Locomotion and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-06-19 paper.tex
Change description of constrained rrt to comply with implementation in hpp-constrained

@@

\begin{algorithm}
\caption{RRT($q_0$)}
\label{algRRT}
\begin{algorithmic}
\STATE $\mathcal{T}.$Init$(q_0)$
\FOR{$i$ = 1 to $K$}
\STATE $q_{rand} \leftarrow $ Rand$(\mathcal{C})$
\STATE $q_{near} \leftarrow $ Nearest$(q_{rand},\mathcal{T})$
\STATE Extend$(q_{near},q_{rand})$
\ENDFOR

@@

  \begin{tikzpicture}[x=0.5cm,y=0.5cm]

@@

    \node [draw,circle,inner sep=2pt] (qn1) at (4.8,2) {};
    \node [draw,circle,inner sep=2pt] (qn2) at (6.2,2) {};
    \node [draw,circle,inner sep=2pt] (qn3) at (7.6,2) {};

    \draw (qn) -- (qn1) -- (qn2) -- (qn3) ;

    \path [<->,thick,red] (3.4,1.5) edge  node [below,black]
          {$\varepsilon$}  (4.8,1.5) ;
          \path [<->,thick,red] (4.8,1.5) edge  node [below,black]
                {$\varepsilon$}  (6.2,1.5) ;
                \path [<->,thick,red] (6.2,1.5) edge  node [below,black]
                      {$\varepsilon$} (7.6,1.5) ;

                      \node [rectangle] (newconf) at (6.2,3.5) {\begin{small}new configurations\end{small}};

                      \draw [->,blue] (newconf) -- (qn1.north) ;
                      \draw [->,blue] (newconf) -- (qn2.north) ;
                      \draw [->,blue] (newconf) -- (qn3.north) ;

                      \node [draw,circle,inner sep=2pt] (qrand) at (13,2) {};
                      \node [below right] at (qrand) {$q_{rand}$};

                      \draw [dashed,thin] (qn3) -- (qrand) ;

@@

  \caption{One step of extension of the RRT algorithm} 
  \label{fig:rrt-extend}
\end{figure}

The equivalent random  extension on a constrained manifold $\mathcal{M}$
starts from a valid configuration $q_{near} \in \mathcal{M}$, and extends the tree
towards a random configuration $q_{rand}$, while keeping the constraints
satisfied. To do so, we add a task to the local planner, whose value
is the distance to  $q_{rand}$ in $\mathcal{C}$. Algorithm~\ref{alg:constrained}
shows the pseudo-code of a constrained random extension from  $q_{near}$ towards
$q_{rand}$ while satisfying the constraints defining $\mathcal{M}$: $Constraints$.
The configuration space task used to explore $\mathcal{C}$ is 
noted $ConfigurationTask$.   This
task is added with the lowest priority. Each new configuration $q_{new}$ generated by the
local planner  is 
checked for constraint satisfaction and collision 
avoidance. If the new configuration is valid, each new edge is checked for collision
avoidance, as in a classic RRT algorithm.

\begin{algorithm}[h]
  \caption{Constrained-Extend($\mathcal{T},q_{near},q_{rand},Constraints$)}
  \label{alg:constrained}
  \begin{algorithmic}
    \STATE $Tasks.$Initialize$(Constraints)$
    \STATE $ConfigurationTask$.SetTarget($q_{rand}$)
    \STATE $Tasks.$Add$(ConfigurationTask)$
    \STATE $q_{current} \leftarrow q_{near}$
    \STATE $State \leftarrow$ Progressing
    \WHILE{$State =  $ Progressing}
    \STATE $q_{new} \leftarrow $ LocalPlannerPerformOneStep($Tasks,q_{current}$)
    \IF{$q_{new} \neq q_{current}$ \\ 
      and CollisionCheck($q_{new}$) = OK\\ 
      and CollisionCheckEdge($q_{current},q_{new}$) = OK \\
      and ConstraintsCheck($q_{new},Constraints$) = OK }
    \STATE $\mathcal{T}$.AddNode($q_{new}$)
    \STATE $\mathcal{T}$.AddEdge($q_{current},q_{new}$)
    \STATE $q_{current} \leftarrow q_{new}$
    \IF{$q_{target} = q_{rand}$}
    \STATE $State \leftarrow$ Reached
    \ENDIF
    \ELSE 
    \STATE $State \leftarrow$ Trapped
    \ENDIF
    \ENDWHILE

Fig. \ref{fig:gikrrt} shows one step of constrained
extension, and the prioritized stack of tasks and constraints solved by the local planner.

\begin{figure}[h]
\centering
  \begin{tikzpicture}[x=0.61cm,y=0.61cm]
    \definecolor{dg}{rgb}{0,0.3,0}
     \draw [dashed,color=dg,thick,text width=2cm] (1,1)
      arc(139:90:10.61) node [above right] {\small{Constrained manifold $\mathcal{M}$}} ; 
      \node [draw,circle,inner sep=2pt,fill=gray] (qnear) at (2,2) {};
      \node [left] (qn0) at (qnear) {$q_{near}$} ;
      \node [draw,circle,inner sep=2pt] (q1) at (3,3) {};
      \node [draw,circle,inner sep=2pt] (q2) at (4,4) {};
      \node [draw,circle,inner sep=2pt] (q3) at (5,5) {};
      \node [draw,circle,inner sep=2pt] (qrand) at (7,7) {};
      \node [below right] (qr0) at (qrand) {$q_{rand}$} ;
      \draw [dashed,thin] (qnear) -- (q1) -- (q2) -- (q3) -- (qrand) ;

    \node [draw,circle,inner sep=2pt,fill=gray] (qc1) at (3.11,2.83) {};
      \draw (qnear) -- (qc1);
      \node [draw,circle,inner sep=2pt,fill=gray] (qc2) at (4.26,3.49) {};
      \draw (qc1) -- (qc2);
     \node [draw,circle,inner sep=2pt,fill=gray] (qc3) at (5.37,3.97) {};
      \draw (qc2) -- (qc3);
      \node [rectangle] (newconf) at (6,2) {\small{new configurations}}; 
      \draw [->,blue] (newconf) -- (qc1) ;
      \draw [->,blue] (newconf) -- (qc2) ;
      \draw [->,blue] (newconf) -- (qc3) ;

      \draw [color=black, thick,fill = white] (15,1) rectangle (17,2);
      \draw [color=black, thick,fill = white] (15,2) rectangle (17,3);
      \draw [color=black, thick,fill = white] (15,3) rectangle (17,5);
      \draw [color=black, thick,fill = white] (15,5) rectangle (17,6);
      \draw [color=black, thick,fill = white] (15,6) rectangle (20,7);
      \draw [color=black, thick,fill = white] (16,1) rectangle (20,6);
      \draw [thick] (16,1) -- (20,1) -- (20,6);
      \draw (16,2) -- (20,2) ;
      \node [text width = 5cm,text centered] at (17.5,6.5) {\large Stack
        of tasks};
      \begin{scriptsize}
        \node at (15.5,5.5) {1};
        \node at (15.5,4) {$\vdots$}; 
        \node at (15.5,2.5) {$n$};
        \node at (15.5,1.5) {$n$+1};

        \node [text width = 2cm,color=dg] at (18.2,4) {\small Constraints defining
          $\mathcal{M}$};
        \node [text width = 3cm,text centered] at (18,1.5)
              { Configuration task towards $q_{rand}$};    
      \end{scriptsize}

\caption{One step of constrained extension, and the stack of tasks used
  to generate new configurations with a local planner starting from $q_{near}$:
  first, the constraints defining $\mathcal{M}$, and at a lower priority
  a configuration task towards $q_{rand}$.}
\label{fig:gikrrt}
\end{figure}

@@

If the animated  trajectory collides with the environment,  we cut the
initial  path   $p$  into  two   sub-paths,  that  we  try   to  animate
recursively. When the  paths to animate are too  short for the robot
nominal  walk parameters, we  accelerate the  steps, and  decrease the
maximum height of  the moving foot. As shown  in previous section, the
walk trajectory  corresponding to  smaller and faster  steps converges
toward the  sliding path.  Algorithm  \ref{alg:walk} shows pseudo-code
that takes  a sliding path $p$  as input and  returns a collision-free
walk trajectory.

>>>>

\usepackage{url}

@@

\begin{algorithm}
\caption{RRT($q_0$)}
\label{algRRT}
\begin{algorithmic}
\STATE $\mathcal{T}.$Init$(q_0)$
\FOR{$i$ = 1 to $K$}
\STATE $q_{rand} \leftarrow $ Rand$(\mathcal{C})$
\STATE $q_{near} \leftarrow $ Nearest$(q_{rand},\mathcal{T})$
\STATE Extend$(\mathcal{T},q_{near},q_{rand})$
\ENDFOR

@@

  \begin{tikzpicture}[x=0.45cm,y=0.45cm]

@@

    \node [draw,circle,inner sep=2pt] (qn2) at (7.6,2) {};
    \node [above left] at (qn2) {$q_{new}$};

    \draw (qn) -- (qn2) ;

    \path[draw=black,line join=miter,line cap=butt,line width=0.800pt, fill=gray]
    (8,2) .. controls (8,4) and (10,6) ..
    (12,5) .. controls (14,4) and (14,4) ..
    (15,2) .. controls (15,0) and (14,0) ..
    (13,0) .. controls (11,1) and (8,1) ..
    (8,2) -- cycle;

    \node [draw,circle,inner sep=2pt,fill=white] (qrand) at (13,2) {};
    \node [below] at (qrand) {$q_{rand}$};

    \node at (10.5,4) {Obstacle};

    \draw [dashed,thin] (qn2) -- (qrand) ;

@@

  \caption{One step of extension of the RRT algorithm. The algorithm tries to add the longest possible edge
  from $q_{near}$ towards $q_{rand}$, while avoiding collisions.} 
  \label{fig:rrt-extend}
\end{figure}

The equivalent random  extension on a constrained manifold $\mathcal{M}$, 
defined by the constraint function $f$,
starts from a valid configuration $q_{near} \in \mathcal{M}$, and extends the tree
towards a random configuration $q_{rand}$, while keeping the constraints
satisfied. Extension attempts orthogonal to $\mathcal{M}$ are useless,
as newly added edges have to be included  in $\mathcal{M}$. To extend in directions
that follow the directions of $\mathcal{M}$, we rely on the framework of 
jacobian-based prioritized inverse kinematics. Algorithm~\ref{alg:constrained} presents 
the adaptation of the classic extend function, and Fig.~\ref{fig:gikrrt} illustrates this
extension. The idea is to first project $q_{rand}$ on the tangent
space to $\mathcal{M}$  at $q_{near}$. Let us call the projected configuration $q_{rand}'$.
$q_{rand}'$ is obtained by one step of prioritized inverse kinematics optimization. Let 
$q_{rand}''$  be the result of a call to $\text{constraint-solver}(q_{rand}',f,\epsilon)$.  It is 
the projection
of $q_{rand}'$ on $\mathcal{M}$. Instead of extending the tree from $q_{near}$ towards 
$q_{rand}$, the algorithm tries to extend from $q_{near}$ towards 
$q_{rand}''$\footnote{The algorithm presented here gives a precise idea of the method, without 
  focusing on technical details. Readers interested in the actual implementation can
  refer to \url{https://github.com/laas/hpp-constrained} and 
  \url{https://github.com/laas/hpp-constrained-planner}, where the open-source 
  code corresponding to this algorithm is made available.}.

\begin{figure}[h]
\centering
\begin{minipage}[c]{0.6\linewidth}
\begin{tikzpicture}[y=0.55pt, x=0.55pt,yscale=-1, inner sep=0pt, outer sep=0pt]
\definecolor{dg}{rgb}{0,0.3,0}
\path[draw=black,line join=miter,line cap=butt,line width=0.800pt, fill=gray]
  (231.3249,143.2249) .. controls (178.0814,153.1889) and (178.5527,172.4247) ..
  (180.8173,200.8036) .. controls (183.2959,231.8645) and (195.6610,188.6038) ..
  (256.5787,230.0980) .. controls (295.8993,256.8813) and (346.4296,308.3911) ..
  (295.9747,178.5802) .. controls (275.5517,126.0356) and (304.8066,129.4735) ..
  (231.3249,143.2249) -- cycle;
\node at (230,180) {Obstacle};
\path[dashed,draw=dg,line join=miter,line cap=butt,line width=1.5pt]
  (88.8934,277.5752) .. controls (196.3490,276.5363) and (532.8532,213.7494) ..
(434.3656,81.6056) node [above = 0.1cm, color=dg, text width=1.8cm] {\small{Constrained manifold $\mathcal{M}$}};
\node [draw,circle,inner sep=2pt,fill = gray] (qn) at (97,277) {};
\node [above = 0.15cm] at (qn) {$q_{near}$};
\node [draw,circle,inner sep=2pt] (qr1) at (367,99) {};
\node [left = 0.2cm] at (qr1) {$q_{rand}$};
\draw [dashed,thin] (qn) -- (450,277);
\node [draw,circle,inner sep=2pt] (qr2) at (367,277) {};
\node [below right = 0.1cm] at (qr2) {$q_{rand}'$};
\node [draw,circle,inner sep=2pt,fill = gray] (qr3) at (345,225) {};
\node [above = 0.22cm] at (qr3) {$q_{rand}''$} ;
\draw [dashed,thin] (qr1) -- (qr2) -- (qr3);
\node [draw,circle,inner sep=2pt,fill = gray] (qnew) at (270,250);
\node [above = 0.1cm, left = 0.2cm] at (qnew) {$q_{new}$};
%\draw [thick] (qn) -- (qnew);

\end{tikzpicture}
\end{minipage}
\begin{minipage}[c]{0.3\linewidth}
\begin{tikzpicture}[x=0.61cm,y=0.61cm]
   \definecolor{dg}{rgb}{0,0.3,0}
      %Stack of tasks:
      \draw [color=black, thick,fill = white] (0,1) rectangle (2,2);
      \draw [color=black, thick,fill = white] (0,2) rectangle (2,3);
      \draw [color=black, thick,fill = white] (0,3) rectangle (2,5);
      \draw [color=black, thick,fill = white] (0,5) rectangle (2,6);
      \draw [color=black, thick,fill = white] (0,6) rectangle (5,7);
      \draw [color=black, thick,fill = white] (1,1) rectangle (5,6);
      \draw [thick] (1,1) -- (5,1) -- (5,6);
      \draw (1,2) -- (5,2) ;
      \node [text width = 5cm,text centered] at (2.5,6.5) {\large Stack
        of tasks};
      \begin{scriptsize}
        \node at (0.5,5.5) {1};
        \node at (0.5,4) {$\vdots$}; 
        \node at (0.5,2.5) {$n$};
        \node at (0.5,1.5) {$n$+1};

        \node [text width = 2cm,color=dg] at (3.2,4) {\small Constraints defining
          $\mathcal{M}$};
        \node [text width = 3cm,text centered] at (3,1.5)
              { Configuration task towards $q_{rand}$};    
      \end{scriptsize}

\end{tikzpicture}
\end{minipage}

\caption{One step of constrained extension, and the stack of tasks used
  to generate $q_{rand}'$ with a prioritized inverse kinematics solver, 
  starting from $q_{near}$:
  first, the constraints defining $\mathcal{M}$, and at a lower priority
  a configuration task towards $q_{rand}$.  $q_{rand}'$ is then projected onto
  $\mathcal{M}$, and a classic RRT extension tries to go as far as possible
  from $q_{near}$ towards $q_{rand}''$.}
\label{fig:gikrrt}
\end{figure}

\begin{algorithm}[h]
  \caption{Constrained-Extend($\mathcal{T},q_{near},q_{rand},f,\epsilon$)}
  \label{alg:constrained}
  \begin{algorithmic}
    \STATE {\color{red} // $I$ denotes the identity matrix.}
    \STATE $q_{rand}' \leftarrow q_{near} +   
    \left(I -  
    \left(\frac{\partial f}{\partial q}(q_{near})\right)^{+}
    \frac{\partial f}{\partial q}(q_{near})\right) (q_{rand} - q_{near})$
    \IF {$\text{constraint-solver}(q_{rand}',f,\epsilon) = success$} 
    \STATE RRT::Extend($\mathcal{T},q_{near},q_{rand}'$)
    \ENDIF
  \end{algorithmic}
\end{algorithm}

@@

If the animated  trajectory collides with the environment,  we cut the
initial  path   $p$  into  two   sub-paths,  that  we  try   to  animate
recursively. When the  paths to animate are too  short for the robot
nominal  walk parameters, we  accelerate the  steps, and  decrease the
maximum height of  the moving foot. As shown  in previous section, the
walk trajectory  corresponding to  smaller and faster  steps converges
toward the  sliding path.  Algorithm ~\ref{alg:walk} shows pseudo-code
that takes  a sliding path $p$  as input and  returns a collision-free
walk trajectory\footnote{The actual implementation of this algorithm is
part of an open-source package, available at
\url{https://github.com/laas/hpp-wholebody-step-planner}.}.
